---
title: "Task 1"
author: "Ganger, Kiara"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r Time Step Formatting, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
df <- read.csv(
  "C:/Users/kiara/OneDrive/WGU MSDA/D213 Advanced Data Analytics/Task 1/teleco_time_series.csv"
)
is.data.frame(df)
df_vector <- df[['Revenue']]
is.vector(df_vector)
any(is.na(df)) # Check for NA values in data set
is.null(df)  # Check for null values in data set
anyDuplicated(df$Day) # Check for duplicates in Day
time_series <- ts(df_vector, frequency = 731 / 24) # Convert to time series
is.ts(time_series) # Verify time series object
```

## Overview

This report analyzes the 90-day revenue forecast using an ARIMA model.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary libraries
library(tseries)
library(tidyverse)  # For base packages and charts
library(forecast)
library(lubridate)  # For date manipulation
library(xts)
```

## Selecting of the ARIMA Model

To select the best ARIMA model, I started by checking the stationarity of the data. The Augmented Dickey-Fuller (ADF) test and the KPSS Test indicated the need for differencing.

```{r Stationarity Test 1, echo=TRUE, message=FALSE, warning=FALSE}
kpss_result <- kpss.test(time_series)
print(kpss_result)# Print the result
adf_test_result <- adf.test(time_series) #Double Checking
print(adf_test_result)# Print the result
```

So I applied a first difference to remove trends and stabilize the time series.

```{r  stabilizing the time series, echo=TRUE, warning=FALSE}
# Apply first-order differencing to the time series
diff_time_series <- diff(time_series, differences = 1)
# Remove NA values introduced by differencing
diff_time_series <- na.omit(diff_time_series)
# Plot the differenced time series to visually check stationarity
plot(
  diff_time_series,
  main = "Differenced Time Series",
  ylab = "Value",
  xlab = "Time",
  col = "darkblue",
  type = "l"
)
```

```{r Train/Test, include=FALSE}
train <- time_series[1:512]
test <- time_series[513:731]
```

After differencing the time, I reviewed the ACF and PACF plots to determine initial values for the AR (p) and MA (q) terms.

```{r ACF and PACF plots, echo=TRUE, paged.print=TRUE}
# Time series plot of the differenced series
plot(
  diff_time_series,
  main = "Differenced Time Series",
  ylab = "Value",
  xlab = "Time",
  col = "darkblue",
  type = "l"
)
# Autocorrelation Function (ACF) plot
acf(
  diff_time_series,
  main = "ACF of Differenced Series",
  col = "darkgreen",
  lwd = 2
)
# Partial Autocorrelation Function (PACF) plot
pacf(
  diff_time_series,
  main = "PACF of Differenced Series",
  col = "darkred",
  lwd = 2
)
```

The plots above provide a value of p=1, d=1, q=0.

From there, I tested various combinations of (p, d, q) parameters to ensure an optimal fit and compared their performance using the Akaike Information Criterion (AIC). The model with the lowest AIC score became the winner giving me the values of p=1, d=1, q=0:

```{r Fitting ARIMA Model}
auto.arima(train, seasonal = T)
fit <- Arima(train, order = c(1, 1, 0), include.drift = TRUE)
fit1 <- arima(train, order = c(1, 1, 1))
fit2 <- arima(train, order = c(1, 2, 0))
fit3 <- arima(train, order = c(0, 1, 0))
AIC(fit, fit1, fit2, fit3)
```

## Prediction Interval of the Forecast

The forecast's prediction interval is 1 day, which makes sense since the dataset is made up of daily revenue over two years. By keeping the interval short and aligned with the granularity of the data, the ARIMA model can better capture seasonality and correlations in revenue patterns. I also generated confidence intervals to show the range where future values are likely to fall. Here’s the code I used for the 90-day forecast and its visualization

```{r}
forecast_90 <- forecast(fit, h = 90)  # Forecast using model
plot(
  forecast_90,
  col = "darkblue",
  main = "90-Day Forecast",
  xlab = "Time",
  ylab = "Revenue"
)
lines(513:(512 + length(test)),
      test,
      col = "darkred",
      lwd = 2)  # Align with test data
legend(
  "topright",
  legend = c("Train", "Test", "Forecast", "95% Confidence Interval"),
  col = c("darkblue", "darkred", "blue", "lightblue"),
  lty = c(1, 1, 1, NA),
  # Line styles (solid for Train/Test/Forecast, none for shaded area)
  lwd = c(2, 2, 2, NA),
  # Line widths for Train/Test/Forecast, none for shaded area
  pch = c(NA, NA, NA, 15),
  # Box for shaded confidence interval
  pt.cex = c(1, 1, 1, 2),
  # Size for shaded confidence interval box
  cex = 0.8
)
```

## Forecast Compared to Test Set

The plot below illustrates the 90-day revenue forecast generated by the ARIMA(1,1,0) model alongside the actual revenue values from the test dataset. This visualization provides a direct comparison of the model's predictions against real-world data, highlighting the model's performance over the test set horizon. The blue line, represents the model's forecasted revenue values for 90 days. The red line is the actual revenue with is set to align with the test set to compare accuracy. The shaded blue area is the Confidence Intervals which is around the forecast illustrates the 95% confidence interval, capturing the range within which the actual values are expected to fall. The ARIMA model was built using only the training data, and its forecast was extended to cover the test set period. The graph highlights how well the model tracks the overall revenue trend while factoring in seasonality and fluctuations in the data. The ARIMA model performs well, demonstrating its ability to capture both the overall upward revenue trend and the seasonal patterns present in the data. The actual values align closely with the forecast, and most fall within the 95% confidence interval. This suggests the model provides a reliable short-term forecast of revenue over the 90-day horizon.

```{r}
# Plot the forecast and test data
plot(
  forecast_90,
  col = "darkblue",
  main = "90-Day Revenue Forecast vs. Actual Data",
  xlab = "Time (Days)",
  ylab = "Revenue ($ Millions)"
)
# Overlay test data for comparison
lines(513:(512 + length(test)), 
      test, 
      col = "darkred", 
      lwd = 2)

# Add a legend for clarity
legend(
  "topright",
  legend = c("Forecast", "Test Data", "95% Confidence Interval"),
  col = c("darkblue", "darkred", "lightblue"),
  lty = c(1, 1, NA),
  lwd = c(2, 2, NA),
  fill = c(NA, NA, "lightblue"),
  border = c(NA, NA, "lightblue"),
  pt.cex = 2,
  cex = 0.8
)

```
